A Bioconductor workflow for processing and analysing spatial proteomics data

### 2018-04-03 bioc-workflow_march2017.tex
started to address comments

<<env, echo=FALSE, warning=FALSE>>=
library("BiocStyle")
suppressPackageStartupMessages(library("knitr"))
suppressPackageStartupMessages(library("xtable"))
suppressPackageStartupMessages(library("gridExtra"))
suppressPackageStartupMessages(library("MSnbase"))
suppressPackageStartupMessages(library("pRoloc"))
suppressPackageStartupMessages(library("pRolocdata"))
cache <- FALSE
@

@@

<<install, eval=FALSE>>=
source("https://bioconductor.org/biocLite.R")
biocLite(c("MSnase", "pRoloc", "pRolocdata", "pRolocGUI"))
@

@@

<<getFilename, cache = cache>>=
library("MSnbase")
extdatadir <- system.file("extdata", package = "pRolocdata")
csvfile <- dir(extdatadir, full.names = TRUE,
          pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
basename(csvfile)
@

@@

<<getEcols>>=
getEcols(csvfile, split = ",", n = 2)
@

@@

<<readMSnSet2>>=
hl <- readMSnSet2(csvfile, ecol = 8:27, fnames = 1, skip = 1)
@

Below, we display a short summary of the data. The data contains 
\Sexpr{nrow(hl)} proteins/features common across the 2 biological replicates
for the respective 2 x 10-plex reporter tags (\Sexpr{ncol(hl)}
columns or samples), along with associated feature meta-data such as
protein markers, protein description, number of quantified peptides
etc (see below).

<<showhl>>=
hl
@

@@

<<assesshl>>=
exprs(hl)[1:5, ]
exprs(hl)[c("Q9ERU9", "Q9Z2R6"), c("X126", "X131.1")]
@

@@

<<removefData>>=
fvarLabels(hl)
fvarLabels(hl)[1:3] <- c("uniprot.accession", "uniprot.id", "description")
fvarLabels(hl)[4:6] <- paste0("peptides.expt", 1:3)
## feature vars 1, 2, and 4 to 6
fData(hl)[1:4, c(1:2, 4:6)]
@

@@

<<pdata>>=
pData(hl)$Replicate <- rep(1:2, each = 10)
pData(hl)$Tag <- sub("\\.1$", "", sub("^X", "", sampleNames(hl)))
pData(hl)
@

@@

<<morepdata>>=
expinfo <- dir(extdatadir, full.names = TRUE,
               pattern = "hyperLOPIT-SIData-fraction-info.csv")

fracinfo <- read.csv(expinfo, row.names=1, skip = 2, 
                     header = FALSE, stringsAsFactors = FALSE)

pData(hl)$Gradient.Fraction <- c(fracinfo[, 1], fracinfo[, 2])
pData(hl)$Iodixonal.Density <- c(fracinfo[, 4], fracinfo[, 5])
pData(hl)
@

@@

<<normsum, eval = FALSE>>= 
hl <- normalise(hl, method = "sum") 
@

@@

<<normcheck>>=
summary(rowSums(exprs(hl[, hl$Replicate == 1])))
summary(rowSums(exprs(hl[, hl$Replicate == 2])))
@

@@

<<reps4combine>>=
library("pRolocdata")
data(hyperLOPIT2015ms3r1)
data(hyperLOPIT2015ms3r2)
@

@@

<<pkgdata, eval=FALSE>>=
pRolocdata()
@

will list the \Sexpr{nrow(pRolocdata()$results)} datasets that are
available in \Biocexptpkg{pRolocdata}.

@@

<<samplenames>>=
identical(sampleNames(hyperLOPIT2015ms3r1), sampleNames(hyperLOPIT2015ms3r2))
hyperLOPIT2015ms3r1 <- updateSampleNames(hyperLOPIT2015ms3r1, 1)
hyperLOPIT2015ms3r2 <- updateSampleNames(hyperLOPIT2015ms3r2, 2)
sampleNames(hyperLOPIT2015ms3r1)
sampleNames(hyperLOPIT2015ms3r2)
@

@@

<<fvarnames>>=
fvarLabels(hyperLOPIT2015ms3r1)
fvarLabels(hyperLOPIT2015ms3r2)
@

@@

<<fvarnames2>>=
fvarLabels(hyperLOPIT2015ms3r1)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r1)[3:5], 1)
hyperLOPIT2015ms3r2 <- updateFvarLabels(hyperLOPIT2015ms3r2, "2", sep = "")
fData(hyperLOPIT2015ms3r1) <- fData(hyperLOPIT2015ms3r1)[1:5]
fData(hyperLOPIT2015ms3r2) <- fData(hyperLOPIT2015ms3r2)[3:5]
fvarLabels(hyperLOPIT2015ms3r1)
fvarLabels(hyperLOPIT2015ms3r2)
@

@@

<<combine, cache = cache>>=
combined <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
combined
@

@@

<<impute, eval=FALSE>>=
hl <- impute(hl, method = "knn")
@

@@

\begin{figure}[!ht]
  \centering
<<namap, out.width=".65\\textwidth">>=
image2(is.na(combined), col = c("black", "white"),
       main = "Missing values (white cells) after combining replicates")
@  
\caption{Heatmap of missing values. Note that the features are
  re-ordered to highlight \hl{clusters} of proteins with similar numbers of
  missing values.}
  \label{fig:namap}
\end{figure}

@@

<<filterNA>>=
combined <- filterNA(combined)
combined
@

@@

\begin{figure}[!ht]
  \centering
<<qcplot, fig.show = "hold", message=FALSE, warning=FALSE, out.width=".45\\textwidth">>=
library("pRoloc")
plot2D(hl, fcol = NULL, col = "black")
plot2D(hl, method = "hexbin")
@  
\caption{PCA plot of the mouse stem cell data \texttt{hl}. Each dot
  represents a single protein, and cluster of proteins represent
  proteins residing in the same sub-cellular niche. The figure on the
  right bins proteins and represent the bins density to highlight the
  presence of protein clusters. }
  \label{fig:pcahl}
\end{figure}

@@

\begin{figure}[!ht]
  \centering
<<qcbx, out.width="\\textwidth", fig.asp=1/2>>=
par(mfrow = c(1, 2), ## creates a two-panel figure
    las = 2,         ## axis labels orientation
    cex.axis = .7)   ## axis label size
o <- order(hl$Iodixonal.Density)
plotDist(hl[, o], pcol = "#00000010", xlab = "")
lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
boxplot(exprs(hl[, o]))
@  
\caption{Protein profiles and distribution of channel intensities. The
  red dots represent the mean relative intensity for each channel. }
  \label{fig:qcbx}
\end{figure}

@@

<<markers>>=
## List available marker sets
pRolocmarkers()
@

@@

<<addmarkers>>=
## Use mouse markers
mrk <- pRolocmarkers(species = "mmus")
head(mrk)

## Add mouse markers
hl <- addMarkers(hl, mrk)
@

@@

\begin{figure}[!ht]
  \centering
<<plotmarkers, out.width="\\textwidth", fig.width=12, fig.asp=1/2>>=
par(mfrow = c(1, 2))
plot2D(hl, main = "pRolocmarkers for mouse")
addLegend(hl, cex = .6)
plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
@  
\caption{Annotated PCA plots of the \texttt{hl} dataset, as produced
  with \texttt{plot2D}.}
  \label{fig:plotmarkers}
\end{figure}

\begin{figure}[!ht]
  \centering
<<plotDist, out.width=".6\\textwidth", fig.asp=3/5>>=
hlo <- hl[, order(hl$Iodixonal.Density)]
plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
         pcol = "purple", fractions = "Tag")
title(main = "Marker occupancy profiles along the gradient")
matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
         lty = 1, col = "darkblue", type = "l")
legend("topleft", c("Mitochondrion", "Peroxisome"),
       lty = 1, col = c("purple", "blue"), bty = "n")
@  
  \caption{Mitochondrion and peroxisome protein profiles.}
  \label{fig:plotDist2}
\end{figure}

@@

\begin{figure}[htb]
  \centering
<<plotmarkers3d, eval=FALSE>>=
plot3D(hl, dims = c(1, 2, 7))
@
  \includegraphics[width=.55\textwidth]{./Figures/plot3d.pdf}
  \caption{Using the \texttt{plot3D} function to visualise the
    \texttt{hl} dataset along PCs 1, 2 and 7. }
  \label{fig:plotmarkers3d}
\end{figure}

@@

<<removeGA>>=
hl <- fDataToUnknown(hl, from = "Golgi apparatus", to = "unknown")
getMarkers(hl)
@

@@

\begin{figure}[!ht]
  \centering
<<foi, fig.width=10, out.width=".8\\textwidth">>=
prots <- c("Q8CG48", "Q8CG47", "Q8K2Z4", "Q8C156")
foi13s <- FeaturesOfInterest(description = "13S consensin proteins",
                             fnames = prots, 
                             object = hl)
foi13s

plot2D(hl)
addLegend(hl, cex = .6)
highlightOnPlot(hl, foi13s)
@
  \caption{Highlighting protein features of interest.}
  \label{fig:foi}
\end{figure}

@@

<<pdtab, echo=FALSE, eval = TRUE, results = 'asis'>>=
i <- c(3, 13)
xtable(pData(hl)[i, ], 
       caption = "Differences in gradient fraction pooling.", 
       label = "tab:pdtab")
@

The more relevant comparison unit is not a single channel, but rather
the complete protein occupancy profiles, which are best visualised 
experiment-wide on a PCA plot. As such, we prefer to focus on the
direct, qualitative comparison of individual replicate PCA plots,
assuring that each displays acceptable sub-cellular resolution. Note
that in the code chunk below, we mirror the x-axis to represent the
two figures with the same orientation. The interactive "compare"
application part of the \Biocpkg{pRolocGUI} package is also useful for
examining replicate experiments (see the next section interactive
visualisation for details).

\begin{figure}[!ht]
  \centering
<<plot2Drep, fig.width=12, out.width="\\textwidth", fig.asp=1/2>>= 
par(mfrow = c(1, 2)) 
plot2D(hl[, hl$Replicate == 1], main = "Replicate 1")
plot2D(hl[, hl$Replicate == 2], main = "Replicate 2", mirrorX = TRUE)
@  
  \caption{PCA plots of replicates 1 and 2}
  \label{fig:plot2Drep}
\end{figure}

In addition, the reproducibility can be assessed by performing
independent classification analyses on each replicate (see the section
on \textit{Supervised machine learning} below) and comparing the the
results. Even when the gradient conditions different (for unexpected
technical or voluntary reasons, to maximise resolution when combining
experiments \cite{Trotter:2010}), one expects agreement in the most
confident organelle assignments.

\clearpage

\section*{Interactive visualisation}

Visualisation and data exploration is an important aspect of data
analyses allowing one to shed light on data structure and patterns of
interest. Using the \Biocpkg{pRolocGUI} package we can interactively
visualise, explore and interrogate quantitative spatial proteomics
data. The \Biocpkg{pRolocGUI} package is currently under active
development and it relies on the \texttt{shiny} framework for
reactivity and interactivity. The package currently distributes 3
different GUI's (\textit{main} (default), \textit{compare} or
\textit{classify}) which are wrapped and launched by the
\texttt{pRolocVis} function.

\subsubsection*{The main application}
In the below code chunk we lauch the main
app (note, we do not need to specify the argument, \texttt{app =
  "main"} as it is the default).

<<mainapp, eval = FALSE, echo = TRUE>>=
library("pRolocGUI")
pRolocVis(hl)
@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{./Figures/mainapp.png}
  \caption{A screen shot of clickable interface and zoomable PCA plot
    of the main app in the \Biocpkg{pRolocGUI} package. }
  \label{fig:app}
\end{figure}

As diplayed in the screenshot in figure \ref{fig:app}, the
\textit{main} application is designed for exploratory data analysis
and is divided into 3 tabs: (1) PCA, (2) Profiles and (3) Table
selection. The default view upon loading is the \textit{PCA} tab,
which features a clickable interface and zoomable PCA plot with an
interactive data table for displaying the quantitation information.
Particular proteins of interest can be highlighted using the text
search box. There is also a \textit{Profiles} tab for visualisation of
the protein profiles, which can be used to examine the patterns of
proteins of interest. The \textit{Table selection} tab provides an
interface to control data table column selection. A short
animation\footnote{\url{https://github.com/lmsimp/bioc-pRoloc-hyperLOPIT-workflow/blob/master/Figures/pRolocVis\_pca.gif}}
illustrating the interface is available in the manuscript
repository~\cite{ghrepo}.

\subsubsection*{The compare application}
The \textit{compare} application is useful for examining two replicate
experiments, or two experiments from different conditions, treatments
etc. The compare application is called by default if the input object
to \texttt{pRolocVis} is an \texttt{MSnSetList} of 2 \texttt{MSnSets},
but it can also be specified by calling the argument \texttt{app =
  "compare"}. For example, in the code chunk below we first create an
\texttt{MSnSetList} of replicates 1 and 2 of the hyperLOPIT data, this
is then passed to \texttt{pRolocVis}.

<<compareapp, eval = FALSE, echo = TRUE>>=
data(hyperLOPIT2015ms3r1)
data(hyperLOPIT2015ms3r2)
hllst <- MSnSetList(list(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))
pRolocVis(hllst, app = "compare")
@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{./Figures/SS_Compare1.jpg}
  \caption{The compare application, main panel}
  \label{fig:compare}
\end{figure}

>>>>

@@

\begin{Schunk}
\begin{Sinput}
> source("https://bioconductor.org/biocLite.R")
> biocLite(c("MSnase", "pRoloc", "pRolocdata", "pRolocGUI"))
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> library("MSnbase")
> extdatadir <- system.file("extdata", package = "pRolocdata")
> csvfile <- dir(extdatadir, full.names = TRUE,
+           pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
> basename(csvfile)
\end{Sinput}
\begin{Soutput}
[1] "hyperLOPIT-SIData-ms3-rep12-intersect.csv.gz"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> getEcols(csvfile, split = ",", n = 2)
\end{Sinput}
\begin{Soutput}
 [1] ""                                  ""                                 
 [3] ""                                  "Experiment 1"                     
 [5] "Experiment 2"                      "Experiment 1"                     
 [7] "Experiment 2"                      "126"                              
 [9] "127N"                              "127C"                             
[11] "128N"                              "128C"                             
[13] "129N"                              "129C"                             
[15] "130N"                              "130C"                             
[17] "131"                               "126"                              
[19] "127N"                              "127C"                             
[21] "128N"                              "128C"                             
[23] "129N"                              "129C"                             
[25] "130N"                              "130C"                             
[27] "131"                               "phenoDisco Input"                 
[29] "phenoDisco Output"                 "Curated phenoDisco Output"        
[31] "SVM marker set"                    "SVM classification"               
[33] "SVM score"                         "SVM classification (top quartile)"
[35] "Final Localization Assignment"     "First localization evidence?"     
[37] "Curated Organelles"                "Cytoskeletal Components"          
[39] "Trafficking Proteins"              "Protein Complexes"                
[41] "Signaling Cascades"                "Oct4 Interactome"                 
[43] "Nanog Interactome"                 "Sox2 Interactome"                 
[45] "Cell Surface Proteins"            
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> hl <- readMSnSet2(csvfile, ecol = 8:27, fnames = 1, skip = 1)
\end{Sinput}
\end{Schunk}

Below, we display a short summary of the data. The data contains 
5032 proteins/features common across the 2 biological replicates
for the respective 2 x 10-plex reporter tags (20
columns or samples), along with associated feature meta-data such as
protein markers, protein description, number of quantified peptides
etc (see below).

\begin{Schunk}
\begin{Sinput}
> hl
\end{Sinput}
\begin{Soutput}
MSnSet (storageMode: lockedEnvironment)
assayData: 5032 features, 20 samples 
  element names: exprs 
protocolData: none
phenoData: none
featureData
  featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
  fvarLabels: X X.1 ... Cell.Surface.Proteins (25 total)
  fvarMetadata: labelDescription
experimentData: use 'experimentData(object)'
Annotation:  
- - - Processing information - - -
 MSnbase version: 2.5.9 
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> exprs(hl)[1:5, ]
\end{Sinput}
\begin{Soutput}
          X126 X127N X127C X128N X128C X129N X129C X130N X130C  X131 X126.1
Q9JHU4   0.028 0.034 0.024 0.014 0.026 0.045 0.107 0.341 0.059 0.321  0.037
Q9QXS1-3 0.039 0.134 0.095 0.053 0.084 0.121 0.107 0.128 0.122 0.117  0.033
Q9ERU9   0.021 0.013 0.014 0.009 0.024 0.054 0.116 0.257 0.209 0.284  0.026
P26039   0.120 0.255 0.148 0.091 0.135 0.095 0.041 0.057 0.014 0.043  0.111
Q8BTM8   0.055 0.139 0.078 0.050 0.077 0.098 0.093 0.171 0.079 0.160  0.062
         X127N.1 X127C.1 X128N.1 X128C.1 X129N.1 X129C.1 X130N.1 X130C.1 X131.1
Q9JHU4     0.064   0.058   0.059   0.067   0.078   0.140   0.208   0.141  0.147
Q9QXS1-3   0.073   0.074   0.062   0.081   0.142   0.190   0.069   0.151  0.125
Q9ERU9     0.017   0.023   0.029   0.039   0.071   0.105   0.171   0.304  0.215
P26039     0.181   0.141   0.144   0.152   0.119   0.075   0.028   0.017  0.033
Q8BTM8     0.108   0.091   0.086   0.099   0.111   0.117   0.095   0.144  0.087
\end{Soutput}
\begin{Sinput}
> exprs(hl)[c("Q9ERU9", "Q9Z2R6"), c("X126", "X131.1")]
\end{Sinput}
\begin{Soutput}
        X126 X131.1
Q9ERU9 0.021  0.215
Q9Z2R6 0.563  0.000
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> fvarLabels(hl)
\end{Sinput}
\begin{Soutput}
 [1] "X"                                 "X.1"                              
 [3] "X.2"                               "Experiment.1"                     
 [5] "Experiment.2"                      "Experiment.1.1"                   
 [7] "Experiment.2.1"                    "phenoDisco.Input"                 
 [9] "phenoDisco.Output"                 "Curated.phenoDisco.Output"        
[11] "SVM.marker.set"                    "SVM.classification"               
[13] "SVM.score"                         "SVM.classification..top.quartile."
[15] "Final.Localization.Assignment"     "First.localization.evidence."     
[17] "Curated.Organelles"                "Cytoskeletal.Components"          
[19] "Trafficking.Proteins"              "Protein.Complexes"                
[21] "Signaling.Cascades"                "Oct4.Interactome"                 
[23] "Nanog.Interactome"                 "Sox2.Interactome"                 
[25] "Cell.Surface.Proteins"            
\end{Soutput}
\begin{Sinput}
> fvarLabels(hl)[1:3] <- c("uniprot.accession", "uniprot.id", "description")
> fvarLabels(hl)[4:6] <- paste0("peptides.expt", 1:3)
> ## feature vars 1, 2, and 4 to 6
> fData(hl)[1:4, c(1:2, 4:6)]
\end{Sinput}
\begin{Soutput}
         uniprot.accession  uniprot.id peptides.expt1 peptides.expt2
Q9JHU4              Q9JHU4 DYHC1_MOUSE            175            166
Q9QXS1-3          Q9QXS1-3  PLEC_MOUSE            123            150
Q9ERU9              Q9ERU9  RBP2_MOUSE            101             90
P26039              P26039  TLN1_MOUSE            101             94
         peptides.expt3
Q9JHU4              322
Q9QXS1-3            174
Q9ERU9              181
P26039              167
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> pData(hl)$Replicate <- rep(1:2, each = 10)
> pData(hl)$Tag <- sub("\\.1$", "", sub("^X", "", sampleNames(hl)))
> pData(hl)
\end{Sinput}
\begin{Soutput}
        Replicate  Tag
X126            1  126
X127N           1 127N
X127C           1 127C
X128N           1 128N
X128C           1 128C
X129N           1 129N
X129C           1 129C
X130N           1 130N
X130C           1 130C
X131            1  131
X126.1          2  126
X127N.1         2 127N
X127C.1         2 127C
X128N.1         2 128N
X128C.1         2 128C
X129N.1         2 129N
X129C.1         2 129C
X130N.1         2 130N
X130C.1         2 130C
X131.1          2  131
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> expinfo <- dir(extdatadir, full.names = TRUE,
+                pattern = "hyperLOPIT-SIData-fraction-info.csv")
> fracinfo <- read.csv(expinfo, row.names=1, skip = 2, 
+                      header = FALSE, stringsAsFactors = FALSE)
> pData(hl)$Gradient.Fraction <- c(fracinfo[, 1], fracinfo[, 2])
> pData(hl)$Iodixonal.Density <- c(fracinfo[, 4], fracinfo[, 5])
> pData(hl)
\end{Sinput}
\begin{Soutput}
        Replicate  Tag Gradient.Fraction Iodixonal.Density
X126            1  126           Cytosol               0.0
X127N           1 127N   1 to 6 (pooled)               6.0
X127C           1 127C   8 to 9 (pooled)              11.0
X128N           1 128N 10 to 11 (pooled)              13.3
X128C           1 128C                12              14.6
X129N           1 129N                14              17.4
X129C           1 129C                16              20.1
X130N           1 130N                18              26.8
X130C           1 130C         Chromatin                NA
X131            1  131                19              34.5
X126.1          2  126           Cytosol               0.0
X127N.1         2 127N   1 to 6 (pooled)               5.2
X127C.1         2 127C   7 to 9 (pooled)              10.0
X128N.1         2 128N 10 to 11 (pooled)              12.5
X128C.1         2 128C                12              14.0
X129N.1         2 129N 14 to 15 (pooled)              17.3
X129C.1         2 129C                17              20.9
X130N.1         2 130N 18 to 19 (pooled)              24.7
X130C.1         2 130C         Chromatin                NA
X131.1          2  131                20              31.9
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> hl <- normalise(hl, method = "sum") 
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> summary(rowSums(exprs(hl[, hl$Replicate == 1])))
\end{Sinput}
\begin{Soutput}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  0.997   0.999   1.000   1.000   1.001   1.003 
\end{Soutput}
\begin{Sinput}
> summary(rowSums(exprs(hl[, hl$Replicate == 2])))
\end{Sinput}
\begin{Soutput}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  0.997   0.999   1.000   1.000   1.001   1.003 
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> library("pRolocdata")
> data(hyperLOPIT2015ms3r1)
> data(hyperLOPIT2015ms3r2)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> pRolocdata()
\end{Sinput}
\end{Schunk}

will list the 75 datasets that are
available in \Biocexptpkg{pRolocdata}.

@@

\begin{Schunk}
\begin{Sinput}
> identical(sampleNames(hyperLOPIT2015ms3r1), sampleNames(hyperLOPIT2015ms3r2))
\end{Sinput}
\begin{Soutput}
[1] TRUE
\end{Soutput}
\begin{Sinput}
> hyperLOPIT2015ms3r1 <- updateSampleNames(hyperLOPIT2015ms3r1, 1)
> hyperLOPIT2015ms3r2 <- updateSampleNames(hyperLOPIT2015ms3r2, 2)
> sampleNames(hyperLOPIT2015ms3r1)
\end{Sinput}
\begin{Soutput}
 [1] "X126.1"  "X127N.1" "X127C.1" "X128N.1" "X128C.1" "X129N.1" "X129C.1"
 [8] "X130N.1" "X130C.1" "X131.1" 
\end{Soutput}
\begin{Sinput}
> sampleNames(hyperLOPIT2015ms3r2)
\end{Sinput}
\begin{Soutput}
 [1] "X126.2"  "X127N.2" "X127C.2" "X128N.2" "X128C.2" "X129N.2" "X129C.2"
 [8] "X130N.2" "X130C.2" "X131.2" 
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> fvarLabels(hyperLOPIT2015ms3r1)
\end{Sinput}
\begin{Soutput}
[1] "EntryName"          "ProteinDescription" "Peptides"          
[4] "PSMs"               "ProteinCoverage"    "markers"           
\end{Soutput}
\begin{Sinput}
> fvarLabels(hyperLOPIT2015ms3r2)
\end{Sinput}
\begin{Soutput}
[1] "EntryName"          "ProteinDescription" "Peptides"          
[4] "PSMs"               "ProteinCoverage"    "markers"           
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> fvarLabels(hyperLOPIT2015ms3r1)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r1)[3:5], 1)
> hyperLOPIT2015ms3r2 <- updateFvarLabels(hyperLOPIT2015ms3r2, "2", sep = "")
> fData(hyperLOPIT2015ms3r1) <- fData(hyperLOPIT2015ms3r1)[1:5]
> fData(hyperLOPIT2015ms3r2) <- fData(hyperLOPIT2015ms3r2)[3:5]
> fvarLabels(hyperLOPIT2015ms3r1)
\end{Sinput}
\begin{Soutput}
[1] "EntryName"          "ProteinDescription" "Peptides1"         
[4] "PSMs1"              "ProteinCoverage1"  
\end{Soutput}
\begin{Sinput}
> fvarLabels(hyperLOPIT2015ms3r2)
\end{Sinput}
\begin{Soutput}
[1] "Peptides2"        "PSMs2"            "ProteinCoverage2"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> combined <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
> combined
\end{Sinput}
\begin{Soutput}
MSnSet (storageMode: lockedEnvironment)
assayData: 6725 features, 20 samples 
  element names: exprs 
protocolData: none
phenoData
  sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
  varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
  varMetadata: labelDescription
featureData
  featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2Y3-3 (6725 total)
  fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
    total)
  fvarMetadata: labelDescription
experimentData: use 'experimentData(object)'
  pubMedIds: 26754106 
Annotation:  
- - - Processing information - - -
Combined [5489,10] and [6268,10] MSnSets Tue Apr  3 12:53:18 2018 
 MSnbase version: 2.5.9 
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> hl <- impute(hl, method = "knn")
\end{Sinput}
\end{Schunk}

@@

\begin{figure}[!ht]
  \centering
\begin{Schunk}
\begin{Sinput}
> image2(is.na(combined), col = c("black", "white"),
+        main = "Missing values (white cells) after combining replicates")
\end{Sinput}
\end{Schunk}
\caption{Heatmap of missing values. Note that the features are
  re-ordered to highlight \hl{clusters} of proteins with similar numbers of
  missing values.}
  \label{fig:namap}
\end{figure}

@@

\begin{Schunk}
\begin{Sinput}
> combined <- filterNA(combined)
> combined
\end{Sinput}
\begin{Soutput}
MSnSet (storageMode: lockedEnvironment)
assayData: 5032 features, 20 samples 
  element names: exprs 
protocolData: none
phenoData
  sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
  varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
  varMetadata: labelDescription
featureData
  featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
  fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
    total)
  fvarMetadata: labelDescription
experimentData: use 'experimentData(object)'
  pubMedIds: 26754106 
Annotation:  
- - - Processing information - - -
Combined [5489,10] and [6268,10] MSnSets Tue Apr  3 12:53:18 2018 
Subset [6725,20][5032,20] Tue Apr  3 12:53:18 2018 
Removed features with more than 0 NAs: Tue Apr  3 12:53:18 2018 
Dropped featureData's levels Tue Apr  3 12:53:18 2018 
 MSnbase version: 2.5.9 
\end{Soutput}
\end{Schunk}

@@

\begin{figure}[!ht]
  \centering
\begin{Schunk}
\begin{Sinput}
> library("pRoloc")
> plot2D(hl, fcol = NULL, col = "black")
> plot2D(hl, method = "hexbin")
\end{Sinput}
\end{Schunk}
\caption{PCA plot of the mouse stem cell data \texttt{hl}. Each dot
  represents a single protein, and cluster of proteins represent
  proteins residing in the same sub-cellular niche. The figure on the
  right bins proteins and represent the bins density to highlight the
  presence of protein clusters. }
  \label{fig:pcahl}
\end{figure}

@@

\begin{figure}[!ht]
  \centering
\begin{Schunk}
\begin{Sinput}
> par(mfrow = c(1, 2), ## creates a two-panel figure
+     las = 2,         ## axis labels orientation
+     cex.axis = .7)   ## axis label size
> o <- order(hl$Iodixonal.Density)
> plotDist(hl[, o], pcol = "#00000010", xlab = "")
> lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
> boxplot(exprs(hl[, o]))
\end{Sinput}
\end{Schunk}
\caption{Protein profiles and distribution of channel intensities. The
  red dots represent the mean relative intensity for each channel. }
  \label{fig:qcbx}
\end{figure}

@@

\begin{Schunk}
\begin{Sinput}
> ## List available marker sets
> pRolocmarkers()
\end{Sinput}
\begin{Soutput}
7 marker lists available:
Arabidopsis thaliana [atha]:
 Ids: TAIR, 543 markers
Drosophila melanogaster [dmel]:
 Ids: Uniprot, 179 markers
Gallus gallus [ggal]:
 Ids: IPI, 102 markers
Homo sapiens [hsap]:
 Ids: Uniprot, 872 markers
Mus musculus [mmus]:
 Ids: Uniprot, 937 markers
Saccharomyces cerevisiae [scer_sgd]:
 Ids: SGD, 259 markers
Saccharomyces cerevisiae [scer_uniprot]:
 Ids: Uniprot Accession, 259 markers
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> ## Use mouse markers
> mrk <- pRolocmarkers(species = "mmus")
> head(mrk)
\end{Sinput}
\begin{Soutput}
                P26039                 Q6PB66                 P11276 
  "Actin cytoskeleton"        "Mitochondrion" "Extracellular matrix" 
                Q6PR54                 Q05793                 P19096 
 "Nucleus - Chromatin" "Extracellular matrix"              "Cytosol" 
\end{Soutput}
\begin{Sinput}
> ## Add mouse markers
> hl <- addMarkers(hl, mrk)
\end{Sinput}
\begin{Soutput}
organelleMarkers
           40S Ribosome            60S Ribosome      Actin cytoskeleton 
                     27                      43                      13 
                Cytosol   Endoplasmic reticulum                Endosome 
                     43                      95                      12 
   Extracellular matrix         Golgi apparatus                Lysosome 
                     10                      27                      33 
          Mitochondrion     Nucleus - Chromatin Nucleus - Non-chromatin 
                    383                      64                      85 
             Peroxisome         Plasma membrane              Proteasome 
                     17                      51                      34 
                unknown 
                   4095 
\end{Soutput}
\end{Schunk}

@@

\begin{figure}[!ht]
  \centering
\begin{Schunk}
\begin{Sinput}
> par(mfrow = c(1, 2))
> plot2D(hl, main = "pRolocmarkers for mouse")
> addLegend(hl, cex = .6)
> plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
\end{Sinput}
\end{Schunk}
\caption{Annotated PCA plots of the \texttt{hl} dataset, as produced
  with \texttt{plot2D}.}
  \label{fig:plotmarkers}
\end{figure}

\begin{figure}[!ht]
  \centering
\begin{Schunk}
\begin{Sinput}
> hlo <- hl[, order(hl$Iodixonal.Density)]
> plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
+          pcol = "purple", fractions = "Tag")
> title(main = "Marker occupancy profiles along the gradient")
> matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
+          lty = 1, col = "darkblue", type = "l")
> legend("topleft", c("Mitochondrion", "Peroxisome"),
+        lty = 1, col = c("purple", "blue"), bty = "n")
\end{Sinput}
\end{Schunk}
  \caption{Mitochondrion and peroxisome protein profiles.}
  \label{fig:plotDist2}
\end{figure}

@@

\begin{figure}[htb]
  \centering
\begin{Schunk}
\begin{Sinput}
> plot3D(hl, dims = c(1, 2, 7))
\end{Sinput}
\end{Schunk}
  \includegraphics[width=.55\textwidth]{./Figures/plot3d.pdf}
  \caption{Using the \texttt{plot3D} function to visualise the
    \texttt{hl} dataset along PCs 1, 2 and 7. }
  \label{fig:plotmarkers3d}
\end{figure}

@@

\begin{Schunk}
\begin{Sinput}
> hl <- fDataToUnknown(hl, from = "Golgi apparatus", to = "unknown")
> getMarkers(hl)
\end{Sinput}
\begin{Soutput}
organelleMarkers
           40S Ribosome            60S Ribosome      Actin cytoskeleton 
                     27                      43                      13 
                Cytosol   Endoplasmic reticulum                Endosome 
                     43                      95                      12 
   Extracellular matrix                Lysosome           Mitochondrion 
                     10                      33                     383 
    Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
                     64                      85                      17 
        Plasma membrane              Proteasome                 unknown 
                     51                      34                    4122 
\end{Soutput}
\end{Schunk}

@@

@@
